---
title: "HW 4 - IE 6600 - Sec 03 - Team 5"
author: "Rahul Dixit, Kirti Aggarwal, Asif Khan"
date: "2/24/2021"
output: pdf_document
---

```{r Set Current Directory, echo=FALSE, warning=FALSE}
# Set the R environment to your current directory
#setwd("put_your_working_directory_having_required_datasets_of_assignment_4")
```

```{r Loading Packages, echo=FALSE, warning=FALSE}
# Loading the required R packages
library(tidyverse)
library(lubridate)
library(corrplot)
library(gridExtra)
library(scales)
library(magrittr)
library(tidyr)
library(ggalluvial)
library(treemapify)
library(wesanderson)
```

```{r Adding datasets, echo=FALSE, warning=FALSE}
# Adding datasets to R
airDelay.df <- read_csv("airlines_delay.csv")
wage.df <- read_csv("wages_jobs.csv")
occupation.df <- read_csv("occupations.csv")
```

## Task 1

### Problem Statement
Create a density plot for all different airlines delays by calling a custom function that has transformed x-axis scale for better visualisation.

### Result
```{r Task 1, echo=FALSE, warning=FALSE}
# Creating a subset of airline delay dataset as defined in the problem statement
delayTypes.df <- airDelay.df %>% 
    select(carrier_delay, 
           late_aircraft_delay, 
           nas_delay, 
           security_delay, 
           weather_delay) %>% 
    drop_na()
delayTypes.df <- gather(delayTypes.df)
# Defining a function to create a density plot 
densityPlotFunc<- function(yourDataFrameName)
{
  densityPlot <- ggplot(yourDataFrameName, 
                        aes(x=value,
                            fill=key)) +
    geom_density(alpha=0.5) +
    scale_x_log10(breaks=trans_breaks("log10", 
                                      function(x) 10^x),
                  labels=trans_format("log10", 
                                      math_format(10^.x))) + 
    ggtitle("Density plot of carrier delays in mins") + 
    xlab("log delay in mins") + 
    ylab("Density")
densityPlot
}
# Using custom function to plot for subset task1
densityPlotFunc(delayTypes.df)                                       
```
### Conclusion

## Task 2
### Problem Statement
Create a correlation plot for selected columns (arr_flights, arr_del15, arr_cancelled, arr_diverted, arr_delay, carrier_delay, weather_delay, nas_delay, security_delay and late_aircraft_delay) having labels properly aligned.

### Result
```{r Task 2, echo=FALSE, warning=FALSE}
# Creating a subset of airline delay dataset as defined in the problem statement
selectedColumnsAirDelay.df <- airDelay.df %>% 
  select(7:16) %>% 
  drop_na()
# Creating a correlation matrix for subset task2
correlationAirDelay <- cor(selectedColumnsAirDelay.df)
# correlation matrix plot using corrplot package
corrplot.mixed(correlationAirDelay,
               lower="number",
               upper="circle",
               lower.col="black",
               tl.pos="d",
               number.cex=0.75,
               tl.cex=0.31)
```
### Conclusion


## Task 3
### Problem statement
Create a heatmap to find difference between male and female employees w.r.t to year and occupation

### Result
```{r Task 3, echo=FALSE, warning=FALSE}
wage_df_selected <- wage.df %>% select(Gender, `Slug PUMS Occupation`, `Total Population`, Year)
wage_df_aggregated <- aggregate(wage_df_selected[,c("Total Population")], 
                by=list(wage_df_selected$`Slug PUMS Occupation`,wage_df_selected$Gender, wage_df_selected$Year), "sum")
wage_df_pivoted <- pivot_wider(data = wage_df_aggregated, names_from = Group.2, values_from =`Total Population`)
wage_df_diff <- wage_df_pivoted %>%
  mutate(Difference = Male-Female) # A negative value indicate more number of female than male employees
ggplot(wage_df_diff, aes(x=Group.3, y=Group.1, fill=Difference)) + 
  geom_tile() +
  scale_y_discrete(labels = function(x) str_wrap(x, width=10)) +
  xlab("Years") +
  ylab("Occupation") +
  ggtitle("Difference between Male and Felmale Employees") +
  
  scale_fill_gradientn(colors=wes_palette( "Zissou1", 100, type = "continuous")) +
  theme_minimal()
```
### Conclusion
As clearly seen from the graph there for the occupation of
1. Flight attendants, there are more female employees than male employees
2. Aircraft mechanics and Reservation Transportation, there are almost same number of male and female employees
3. Aircraft pilots, there are more male employees than female employees

## Task 4
### Problem statement
Create an Alluvial Chart for Occupation and Year w.r.t Population, showing inter-relation of Gender.

```{r Task 4, echo=FALSE, warning=FALSE}
ggplot(as.data.frame(wage.df), aes(y =`Total Population`, axis1 = `PUMS Occupation`, axis2 = Year)) +
  geom_alluvium(aes(fill = Gender), width = 1/12) +
  geom_stratum(width = 1/4) +
  scale_x_discrete(limits = c("Occupation", "Year"), expand = c(.1, 0)) +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)), size =1.45, min.y=100) +
  theme_minimal()
```
### Conclusion
1. There are more male employees from the year 2014 to 2018 having an occupation in Aircraft Mechanics, Pilots.
2. There are more female employees from the year 2014 to 2018 having an occupation in Customer service, Flight Attendants and reservation transportation.


## Task 5
### Problem statement
Create a stacked bar plot of Occupation vs Wage (in year 2018) on the basis of Gender as different color

### Result
```{r Task 5, echo=FALSE, warning=FALSE}
# Creating a subset of occupation dataset as defined in the problem statement
wageGenderOccupation.df <- wage.df %>% 
  filter(Year == 2018) %>% 
  select(`PUMS Occupation`, 
         `Average Wage`, 
         Gender) %>% 
  group_by(`PUMS Occupation`)
wageGenderOccupation.df$`Average Wage` <- round(wageGenderOccupation.df$`Average Wage`, digits = 2)
# stacked bar plot using ggplot package 
ggplot(wageGenderOccupation.df, aes(fill=Gender, 
                  y=`Average Wage`, 
                  x=reorder(`PUMS Occupation`,-`Average Wage`))) +
  geom_bar(stat="identity") +
  labs(x = "Occupation", 
       y = "Average Wage", 
       title = "Stacked Bar Plot of Gender on basis of Wages vs Occupation") +
  theme_minimal(base_size = 8) +
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1))
    

```
### Conclusion
1. Average wage for different occupation is more for Males as compared to females for 3 categories (Aircraft Pilots and Flight engineers, Aircraft Mechanics and service technicians Flight Attendants) out of 5. Whereas 2 categories have comparable wage for both males and females.
2. Aircraft Pilots and Flight engineers has the highest average wage.
3. Customer services representatives has the lowest average wage. 


```{r Task 6, echo=FALSE, warning=FALSE}
# Creating a subset of occupation dataset as defined in the problem statement
occupdationGroupAndDetails.df <- occupation.df %>% 
  group_by(`Detailed Occupation`, 
           `Major Occupation Group`) %>% 
  summarise(Workforce=sum(`Total Population`, 
                          na.rm = TRUE), 
            .groups = "drop")
# tree map plot using ggplot and treemapify packages
ggplot(occupdationGroupAndDetails.df, aes(area=Workforce,
                  fill=`Major Occupation Group`,
                  label=`Detailed Occupation`,
                  subgroup=`Major Occupation Group`)) + 
  geom_treemap() +
  labs(title = "Workforce Distribution by Detailed Workforce 2018") +
  geom_treemap_text(fontface = "italic",
                        colour = "white",
                        place = "centre",
                        grow = F,
                        reflow = T) +
  theme(legend.title = element_text(size = 5), 
        legend.text = element_text(size = 5))
```

### Conclusion
1. Service Occupation is the subgroup where maximum number of people work.
2. Natural Resources, construction and maintenance occupation is the subgroup where least number of people work.