---
title: "HW 4 - IE 6600 - Sec 03 - Team 5"
author: "Rahul Dixit, Kirti Aggarwal, Asif Khan"
date: "2/19/2021"
output: pdf_document
---

```{r Set Current Directory, echo=FALSE, warning=FALSE}
# Set the R environment to your current directory
#setwd("/Users/kirtiagwl1595/Desktop/Computation and Data Visualisation/Assignment/Assignment 4")
#setwd("/Users/asif/NEU/IE-6600 - Comp Viz/Assignment/ComputerVizAssignments/HW4")
```

```{r Loading Packages, echo=FALSE, warning=FALSE}}
# Loading the required R packages
library(tidyverse)
library(lubridate)
library(corrplot)
library(gridExtra)
library(scales)
library(reshape2)
library(magrittr)
library(tidyr)
library(ggplot2)
library(dplyr)
library(corrr)
library(ggalluvial)
library(treemapify)
```

```{r Adding datasets, echo=FALSE, warning=FALSE}
# Adding datasets to R
airDelay.df <- read_csv("airlines_delay.csv")
wage.df <- read_csv("wages_jobs.csv")
occupation.df <- read_csv("occupations.csv")
```

## Task 1

### Problem Statement
Create a density plot for all different airlines delays by calling a custom function that has transformed x-axis scale for better visualisation.

### Result
```{r Task 1, echo=FALSE, warning=FALSE}
# Creating a subset of airline delay dataset as defined in the problem statement
delayTypes.df <- airDelay.df %>% 
    select(carrier_delay, 
           late_aircraft_delay, 
           nas_delay, 
           security_delay, 
           weather_delay) %>% 
    drop_na()
delayTypes.df <- gather(delayTypes.df)
# Defining a function to create a density plot 
densityPlotFunc<- function(yourDataFrameName)
{
  densityPlot <- ggplot(yourDataFrameName, 
                        aes(x=value,
                            fill=key)) +
    geom_density(alpha=0.5) +
    scale_x_log10(breaks=trans_breaks("log10", 
                                      function(x) 10^x),
                  labels=trans_format("log10", 
                                      math_format(10^.x))) + 
    ggtitle("Density plot of carrier delays in mins") + 
    xlab("log delay in mins") + 
    ylab("Density")
densityPlot
}
# Using custom function to plot for subset task1
densityPlotFunc(delayTypes.df)                                       
```
### Conclusion

## Task 2
### Problem Statement
Create a correlation plot for selected columns (arr_flights, arr_del15, arr_cancelled, arr_diverted, arr_delay, carrier_delay, weather_delay, nas_delay, security_delay and late_aircraft_delay) having labels properly aligned.

### Result
```{r Task 2, echo=FALSE, warning=FALSE}
# Creating a subset of airline delay dataset as defined in the problem statement
selectedColumnsAirDelay.df <- airDelay.df %>% 
  select(7:16) %>% 
  drop_na()
# Creating a correlation matrix for subset task2
correlationAirDelay <- cor(selectedColumnsAirDelay.df)
# correlation matrix plot using corrplot package
corrplot.mixed(correlationAirDelay,
               lower="number",
               upper="circle",
               lower.col="black",
               tl.pos="d",
               number.cex=0.75,
               tl.cex=0.31)
```
### Conclusion


## Task 3
### Problem statement
### From wages_jobs.csv generate a heat map as shown below. The variable Difference is defined as the difference between number of male employees and the number of female employees. A negative value indicate more number of female than male employees. In addition to the plot critique the below visualization:

### • Transform the x axis using a log10 to better visualize the data spread.

```{r}
library(wesanderson)
wage_df_selected <- wage.df %>% select(Gender, `Slug PUMS Occupation`, `Total Population`, Year)
wage_df_aggregated <- aggregate(wage_df_selected[,c("Total Population")], 
                by=list(wage_df_selected$`Slug PUMS Occupation`,wage_df_selected$Gender, wage_df_selected$Year), "sum")
wage_df_pivoted <- pivot_wider(data = wage_df_aggregated, names_from = Group.2, values_from =`Total Population`)
wage_df_diff <- wage_df_pivoted %>%
  mutate(Difference = Male-Female) # A negative value indicate more number of female than male employees
ggplot(wage_df_diff, aes(x=Group.3, y=Group.1, fill=Difference)) + 
  geom_tile() +
  scale_y_discrete(labels = function(x) str_wrap(x, width=10)) +
  xlab("Years") +
  ylab("Occupation") +
  ggtitle("Difference between Male and Felmale Employees") +
  scale_fill_gradientn(colors=wes_palette( "Zissou1", 100, type = "continuous")) +
  theme_minimal()
```


## Task 4
### Problem statement
### From wages_jobs.csv generate the following alluvial chart.

```{r}
ggplot(as.data.frame(wage.df), aes(y =`Total Population`, axis1 = `PUMS Occupation`, axis2 = Year)) +
  geom_alluvium(aes(fill = Gender), width = 1/12) +
  geom_stratum(width = 1/4) +
  scale_x_discrete(limits = c("Occupation", "Gender"), expand = c(.1, 0)) +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)), size =1.45, min.y=100) +
  theme_minimal()
```
### Conclusion


## Task 5
### Problem statement
Create a stacked bar plot of Occupation vs Wage (in year 2018) on the basis of Gender as different color

### Result
```{r Task 5, echo=FALSE, warning=FALSE}
# Creating a subset of occupation dataset as defined in the problem statement
wageGenderOccupation.df <- wage.df %>% 
  filter(Year == 2018) %>% 
  select(`PUMS Occupation`, 
         `Average Wage`, 
         Gender) %>% 
  group_by(`PUMS Occupation`)
wageGenderOccupation.df$`Average Wage` <- round(wageGenderOccupation.df$`Average Wage`, digits = 2)
# stacked bar plot using ggplot package 
ggplot(wageGenderOccupation.df, aes(fill=Gender, 
                  y=`Average Wage`, 
                  x=reorder(`PUMS Occupation`,-`Average Wage`))) +
  geom_bar(stat="identity", position = "fill") +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(x = "Occupation", 
       y = "Average Wage", 
       title = "Stacked Bar Plot of Gender on basis of Wages vs Occupation") +
  theme_minimal(base_size = 8) +
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1))
    

```
### Conclusion
1. Average wage for different occupation is more for Males as compared to females for 3 categories (Aircraft Pilots and Flight engineers, Aircraft Mechanics and service technicians Flight Attendants) out of 5. Whereas 2 categories have comparable wage for both males and females.
2. Aircraft Pilots and Flight engineers has the highest average wage.
3. Customer services representatives has the lowest average wage. 

```{r}
wage_jobs<-read.csv("wages_jobs.csv", na.strings = "")
df1<-wage_jobs %>% filter(Year==2018) %>% group_by(PUMS.Occupation,Gender) %>% summarise(avg_wage=sum(Average.Wage))%>%
  mutate(percent=(avg_wage/sum(avg_wage))*100) %>% drop_na()
df1$avg_wage<-round(df1$avg_wage,digits = 2)
task1<- ggplot(df1,aes(y=PUMS.Occupation,percent , fill=Gender))+geom_bar(stat="identity",width=0.5)+geom_text(aes(label=avg_wage),position = position_stack(vjust=0.5, reverse = T))+
  theme(axis.text.y = element_text(size=6)) + xlab("Average wage in percentage")+ylab("Occupation")
task1
```


```{r Task 6, echo=FALSE, warning=FALSE}
# Creating a subset of occupation dataset as defined in the problem statement
occupdationGroupAndDetails.df <- occupation.df %>% 
  group_by(`Detailed Occupation`, 
           `Major Occupation Group`) %>% 
  summarise(Workforce=sum(`Total Population`, 
                          na.rm = TRUE), 
            .groups = "drop")
# tree map plot using ggplot and treemapify packages
ggplot(occupdationGroupAndDetails.df, aes(area=Workforce,
                  fill=`Major Occupation Group`,
                  label=`Detailed Occupation`,
                  subgroup=`Major Occupation Group`)) + 
  geom_treemap() +
  labs(title = "Workforce Distribution by Detailed Workforce 2018") +
  geom_treemap_text(fontface = "italic",
                        colour = "white",
                        place = "centre",
                        grow = F,
                        reflow = T) 
```